Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-02-23 local-capability-machine.tex
added lemma about rwlx and pwl regions and used it in g1 proof

            \item $\memSat[n_6-l]{\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0}{W_{10}}$ \label{g1:memsatw10}\\
              First notice that from %TODO write more about why this is the case. Lemma?
              \begin{itemize}
              \item $\npair[n_4]{\reg_3} \in \stdrr(W_3)$
              \item $\memSat[n_4]{\ms_3}{W_3}$
              \item $\reg(r_\stk).\perm = \rwlx$
              \end{itemize}
              we get that there exists a region, $r_{\adv \stk}$ in $W_3$ such that
              \[
                W_3(r_{\adv \stk}) = \iota_{\stk_a,\stk_b}^\pwl
              \]
              and $\dom(\ms_\unused) \subseteq [\stk_a,\stk_b]$. Now take $\ms_{\adv \stk} = \ms_r|_{[\stk_a,\stk_b]}$ (notice this not all of $[\stk_a,\stk_b]$ is in the domain of $\ms_r$).
              We know
              \begin{equation}
                \label{eq:g1:ms7heapsat}
                \memSat[n_6]{\ms_7}{\revokeTemp{W_8}}
              \end{equation}
              and
              \begin{equation}
                \label{eq:g1:ms3heapsat}
                \memSat[n_4]{\ms_3}{W_3}
              \end{equation}
              which gives us two partitions say $P_8$ and $P_3$ respectively. Now define the partition $P$ as follows:
              \[
                P(r) =
                \begin{cases}
                  P_8(r) & r \in \dom(\erase{W_8}{\perma}) \\
                  \ms_{\adv \stk} \uplus \ms_0 & r = r_{\adv \stk}\\
                  P_3(r) & \text{otherwise}
                \end{cases}
              \]
              Now let $r \in \active{W}$, $n_7 < n_6 - l$, and $W(r) = (\_,s,\_,\_,H)$ and show
              \[
                \npair[n_7]{P(r)} \in H(s)(\xi^{-1}(W_{10})).
              \]
              Consider the following cases
              \begin{enumproof}
                \item $r \in \dom(\erase{W_8}{\perma})$ \\
                  Use \ref{eq:g1:ms7heapsat}, the fact that $W_{10} \futurestr \revokeTemp{W_8}$ and that permanent regions respect future private world.
                \item $r = r_{\adv \stk}$ \\
                  In this case we know the region is $\iota^\pwl_{\stk_a,\stk_b}$, so we need to show
                  \[
                    \npair[n_7]{\ms_{\adv \stk} \uplus \ms_0} \in H^\pwl_{\stk_a,\stk_b}(1)(\xi^{-1}(W_{10}))
                  \]
                  which amounts to showing
                  \[
                    \forall a \in \dom(\ms_0) \ldotp  \npair[n_7-1]{\ms_0(a)} \in \stdvr(W_{10}),
                  \]
                  which is trivial, and
                  \[
                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_7-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_{10})
                  \]
                  here we use that \ref{eq:g1:ms3heapsat} entails
                  \[
                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_4-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_3)
                  \]
                  and the fact that $\stdvr$ is monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$, and $\stdvr(W_{10})$ is downwards-closed.
                \item otherwise \\
                  Use \ref{eq:g1:ms3heapsat}, $W_{10} \futurewk W_3$, and the fact that for a temporary region $H(s)$ is monotone w.r.t. $\futurewk$.
              \end{enumproof}

@@

@@

\begin{lemma}
  \label{lem:revoketemp-writecond}
  If
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  then
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
  \[
    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
  \]
  and 
  \begin{equation*}
    \revokeTemp(W)(r) \text{ is address-stratified}
  \end{equation*}
  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
\end{proof}

>>>>

            \item $\memSat[n_6-l]{\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0}{W_{10}}$ \label{g1:memsatw10}\\
              First notice that from %TODO write more about why this is the case. Lemma?
              \begin{itemize}
              \item $\npair[n_4]{\reg_3} \in \stdrr(W_3)$
              \item $\memSat[n_4]{\ms_3}{W_3}$
              \item $\reg(r_\stk).\perm = \rwlx$
              \end{itemize}
              using Lemma~\ref{lem:pwl-stack} we get that there exists a region, $r_{\adv \stk}$ such that
              \[
                W_3(r_{\adv \stk}) \nequal \iota_{\stk_a,\stk_b}^\pwl
              \]
              and $\dom(\ms_\unused) \subseteq [\stk_a,\stk_b]$. Now take $\ms_{\adv \stk} = \ms_r|_{[\stk_a,\stk_b]}$ (notice this not all of $[\stk_a,\stk_b]$ is in the domain of $\ms_r$).
              We know
              \begin{equation}
                \label{eq:g1:ms7heapsat}
                \memSat[n_6]{\ms_7}{\revokeTemp{W_8}}
              \end{equation}
              and
              \begin{equation}
                \label{eq:g1:ms3heapsat}
                \memSat[n_4]{\ms_3}{W_3}
              \end{equation}
              which gives us two partitions say $P_8$ and $P_3$ respectively. Now define the partition $P$ as follows:
              \[
                P(r) =
                \begin{cases}
                  P_8(r) & r \in \dom(\erase{W_8}{\perma}) \\
                  \ms_{\adv \stk} \uplus \ms_0 & r = r_{\adv \stk}\\
                  P_3(r) & \text{otherwise}
                \end{cases}
              \]
              Now let $r \in \active{W}$, $n_7 < n_6 - l$, and $W(r) = (\_,s,\_,\_,H)$ and show
              \[
                \npair[n_7]{P(r)} \in H(s)(\xi^{-1}(W_{10})).
              \]
              Consider the following cases
              \begin{enumproof}
                \item $r \in \dom(\erase{W_8}{\perma})$ \\
                  Use \ref{eq:g1:ms7heapsat}, the fact that $W_{10} \futurestr \revokeTemp{W_8}$ and that permanent regions respect future private world.
                \item $r = r_{\adv \stk}$ \\
                  In this case we know the region is $\iota^\pwl_{\stk_a,\stk_b}$, so we need to show
                  \[
                    \npair[n_7]{\ms_{\adv \stk} \uplus \ms_0} \in H^\pwl_{\stk_a,\stk_b}(1)(\xi^{-1}(W_{10}))
                  \]
                  which amounts to showing
                  \[
                    \forall a \in \dom(\ms_0) \ldotp  \npair[n_7-1]{\ms_0(a)} \in \stdvr(W_{10}),
                  \]
                  which is trivial, and
                  \[
                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_7-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_{10})
                  \]
                  here we use that \ref{eq:g1:ms3heapsat} entails
                  \[
                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_4-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_3)
                  \]
                  and the fact that $\stdvr$ is monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$, and $\stdvr(W_{10})$ is downwards-closed.
                \item otherwise \\
                  Use \ref{eq:g1:ms3heapsat}, $W_{10} \futurewk W_3$, and the fact that for a temporary region $H(s)$ is monotone w.r.t. $\futurewk$.
              \end{enumproof}

@@

\begin{lemma}
  \label{lem:pwl-stack}
  For all $W$, $\start$, $\addrend$, $n$, $\ms$ if
  \begin{itemize}
  \item $\memSat{\ms}{W}$
  \item $\npair{((\rwlx,\gl),\start,\addrend,\addr)} \in \stdvr(W)$
  \item $\start \leq \addrend$
  \end{itemize}
  then
  \[
    \exists r, \start',\addrend' \ldotp [\start,\addrend] \subseteq [\start',\addrend'] \land W(\var{r}) \nequal \iota^\pwl_{\start',\addrend'}
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:pwl-stack}]
  Assume
  \begin{enumproof}
    \item $\npair{((\rwlx, \gl),b,e,a)} \in \stdvr(W)$ \label{ass:pwl-stack:val-rel}
    \item $\memSat{\ms}{W}$ \label{ass:pwl-stack:mem-sat}
  \end{enumproof}
  From Assumption~\ref{ass:pwl-stack:val-rel}, we get $r_1$, $r_2$, $b_1$, $b_2$, $e_1$ and $e_2$ such that
  \begin{enumproof}[resume]
    \item $r_1 \in \localityReg(\gl,W)$
    \item $r_2 \in \localityReg(\gl,W)$
    \item $[b,e] \subseteq [b_1,e_1]$\label{ass:pwl-stack:b1}
    \item $[b,e] \subseteq [b_2,e_2]$ \label{ass:pwl-stack:b2}
    \item $W(r_1) \nsubsim \iota^\pwl_{b_1,e_1}$ \label{ass:pwl-stack:subsim}
    \item $W(r_2) \nsupsim \iota^\pwl_{b_2,e_2}$ \label{ass:pwl-stack:supsim}
    \item $W(r_2)$ is address-stratified. \label{ass:pwl-stack:add-stra}
  \end{enumproof}
From Assumption~\ref{ass:pwl-stack:mem-sat}, we get partitionen $P$ s.t.
\[
  \memSat[n,p]{\ms}{W}
\]
Say $P(r_1) = \ms_1$ and $P(r_2) = \ms_2$. First from $\npair{\ms_1} \in W(r_1).H\;W(r_1).s\;\xi^{(-1)}(W)$ using \label{ass:pwl-stack:subsim}, we get $\npair{\ms_1} \in H^\pwl_{b_1,e_1}\;1\;\xi^{(-1)}(W)$ which means $\dom(\ms_1) = [b_1,e_1]$.

Second we know $\npair{[b_2 \mapsto 0, \dots , e_2 \mapsto 0]} \in H^\pwl_{b_2,e_2}\;1\;\xi^{(-1)}(W)$ and $\npair{\ms_2} \in W(r_2).H\;W(r_2).s\;\xi^{(-1)}(W)$ which by Assumption~\ref{ass:pwl-stack:supsim} and \ref{ass:pwl-stack:add-stra} means $\dom(\ms_2) = [b_2,e_2]$.

Now assume for contradition $r_1 \neq r_2$, then we have a contradiction with $\memSat[n,p]{\ms}{W}$ because $\ms_1$ and $\ms_2$ are not disjoint (by Assumptions~\ref{ass:pwl-stack:b1} and \ref{ass:pwl-stack:b2}). So $r_1=r_2$ which also means $[b_1,e_1] = [b_2,e_2]$, so from Assumption~\ref{ass:pwl-stack:subsim} and \ref{ass:pwl-stack:supsim}, we get $W(r_1) \nsim \iota^\pwl_{b_1,e_1}$ which by Lemma~\ref{lem:pwl-nsim-view} means $W(r_1) \nequal \iota^\pwl_{b_1,e_1}$
\end{proof}

@@

\begin{lemma}
  \label{lem:revoketemp-writecond}
  If
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  then
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
  \[
    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
  \]
  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
  \[
    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
  \]
  and 
  \begin{equation*}
    \revokeTemp{W}(r) \text{ is address-stratified}
  \end{equation*}
  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
\end{proof}
